---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

library(simts)
```


# `simts` R Package

This repository holds the Time Series Simulation (simts) R package.  

Below are examples of the capabilities of the `simts` package.




```{r}
datasets::Nile
```




```{r, fig.height = 5, fig.width = 7.25}
n = 1000
model = ARMA(ar = c(0.9, -0.5), ma = 0.3, sigma2 = 1)
Xt = gen_gts(n = n, model  = model)
plot(Xt)

model = ARMA(ar = c(0.9, -0.5), ma = 0.3, sigma2 = 1) + WN(sigma = 4)
Yt = gen_gts(n = n, model  = model)
plot(Yt)
```

```{r, fig.height = 5, fig.width = 7.25}
model = SARMA(ar = 0, ma = 0.3, sar = 0.9, sma = 0, s = 10, sigma2 = 1)
Xt = gen_gts(n = n, model  = model)
plot(Xt)
```


# Install Instructions
To install the `simts` package, there is currently one option: GitHub (Developmental).

## Recommended R Interface

We firmly recommend that any users of this package use the [RStudio IDE](https://www.rstudio.com/products/rstudio/download/) over the default R GUI.

## Installing the package through GitHub (Developmental)

For users who are interested in having the latest and greatest developments withing wavelets or simts methodology, this option is ideal. Though, there is considerably more work that a user must do to have a stable version of the package. **The setup to obtain the development version is platform dependent.**

Specifically, one **must** have a compiler installed on your system that is compatible with R.

For help on obtaining a compiler consult:

* [OS X](http://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-os-x/)
* [Windows](https://cran.r-project.org/bin/windows/Rtools/)

Depending on your operating system, further requirements exist such as:

**OS X**

Some user report the need to use X11 to suppress shared library errors. To install X11, visit [xquartz.org](http://www.xquartz.org/)

**Linux**

Both curl and libxml are required.

For **Debian** systems, enter the following in terminal:

```{r, eval = F, engine='bash'}
sudo apt-get install curl libcurl3 libcurl3-dev libxml2 libxml2-dev
```

For **RHEL** systems, enter the following in terminal:

```{r, eval = F, engine='bash'}
sudo yum install curl curl-devel libxml2 libxml2-dev
```

**All Systems**

With the system dependency taken care of, we continue on by installing the R specific package dependencies and finally the package itself by doing the following in an R session: 

```{r, eval = F}
# Install dependencies
install.packages(c("RcppArmadillo","ggplot2","reshape2","devtools","knitr","rmarkdown"))

# Install the package from GitHub without Vignettes/User Guides
devtools::install_github("SMAC-Group/simts")
```

## Supplementary data package

To test the package performance on real-world data that is *stationary* or work with some of the examples, you will need to download and install the `imudata` and/or the `datapkg` R package.

To do so, please use the following installation method within the `simts` R package:

```{r, eval = F}
# Install the imudata package containing real world IMU data sets
simts::install_imudata()

# Install the datapkg package containing miscellaneous data sets
simts::install_datapkg()
```

For more information about the `imudata` and `datapkg` package, see the <https://github.com/SMAC-Group/imudata> and <https://github.com/SMAC-Group/datapkg>.

