---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

library(simts)
```

[![Travis-CI Build Status](https://travis-ci.org/SMAC-Group/simts.svg?branch=master)](https://travis-ci.org/SMAC-Group/simts)

# `simts` R Package

This repository holds the Time Series Simulation (simts) R package. `simts` generates various time series objects for use in other packages.

To see what `simts` is capable of, please refer to the vignettes.

# Install Instructions

To install the `simts` package, there is currently one option: GitHub (Developmental).

## Installing the package through GitHub (Developmental)

For users who are interested in having the latest and greatest developments in wavelets and simts methodology, this option is ideal. Though, more dependancies are required to run a stable version of the package. Most importantly, users **must** have a compiler installed on their machine that is compatible with R (e.g. Clang).

*The setup to obtain the development version of `simts` is platform dependent.*

### Requirements and Dependancies

**OS X**

Some users report the need to use X11 to suppress shared library errors. To install X11, visit [xquartz.org](http://www.xquartz.org/).

**Linux**

Both curl and libxml are required.

For **Debian** systems, enter the following in terminal:

```{r, eval = F, engine='bash'}
sudo apt-get install curl libcurl3 libcurl3-dev libxml2 libxml2-dev
```

For **RHEL** systems, enter the following in terminal:

```{r, eval = F, engine='bash'}
sudo yum install curl curl-devel libxml2 libxml2-dev
```

**All Systems**

The following R packages are also required. If you have made it this far, run the following code in an R session and you will be ready to use the devlopment version of `simts`.

```{r, eval = F}
# Install dependencies
install.packages(c("RcppArmadillo","devtools","knitr","rmarkdown"))

# Install the package from GitHub without Vignettes/User Guides
devtools::install_github("SMAC-Group/simts")
```


Please report any bugs to contact@smac-group.com
