---
author: "Justin Lee"
title: "Simulations from Time Series Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulations from Time Series Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
library(simts)
library(datasets)
```

In this section, we will breifly explain demo `simts` functionality by performing simulations with time series data. In the examples shown here, time series data may take the form of either a one-column `matrix`, `data.frame`, or numeric `vector`.

The following three datasets stored in `simts` will be used as examples.

* [hydro](https://rdrr.io/github/SMAC-Group/datapkg/man/hydro.html): Hydrology data that indicates a robust approach may be preferred to a classical approach when estimating time series. Time series with frequency 12 starting at 1907 and going to 1972 for total of 781 observations.
* [savingrt](https://rdrr.io/github/SMAC-Group/datapkg/man/savingrt.html): Personal saving as a percentage of disposable personal income (DPI), frequently referred to as "the personal saving rate," is calculated as the ratio of personal saving to DPI. GTS time series with frequency 12 starting in year 1959 and going to 2016 for a total of 691 observations.
* [Nile](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/Nile.html): Measurements of the annual flow of the river Nile at Aswan (formerly Assuan), 1871–1970, in $10^8 m^3$, “with apparent changepoint near 1898” (Cobb, 1978, Table 1, p.249).

## Generate Simulations 

The code below shows how to setup a time series oriented object with the `gts()` function. Here we take samples from each dataset at a rate of `freq` ticks per sample. By applying the `plot()` on the result of a `gts()` simulation, we can observe a visualization of the data simulated. 

### Hydrology 

```{r}
# Load hydro dataset
data("hydro")
# Simulate based on data
hydro = gts(as.vector(hydro), start = 1907, freq = 12, unit_ts = "in.", 
            name_ts = "Precipitation", data_name = "Hydrology data")
# Simulate latent time series based on data 
hydro = lts(as.vector(hydro), start = 1907, freq = 12, unit_ts = "in.", 
            name_ts = "Precipitation", data_name = "Hydrology data")


# Load savingrt dataset
data("savingrt")
# Simulate based on data
savingrt = gts(as.vector(savingrt), start = 1959, freq = 12, unit_ts = "%", 
            name_ts = "Saving Rates", data_name = "US Personal Saving Rates")

# Load Nile dataset
Nile = datasets::Nile
nile = gts(as.vector(Nile), start = 1871, end = 1970, freq = 1, 
           unit_ts = bquote(paste(10^8," ",m^3)), name_ts = "Flow", 
           name_time = "Years", data_name = "Annual Flow of river Nile")
```


## Visualizations 


```{r, fig.align='center', fig.height = 5, fig.width = 7.25}
# Plot hydro simulation
plot(hydro)

# Plot savingrt simulation
plot(savingrt)

# Plot Nile simulation 
plot(nile)
```