---
author: "Justin Lee"
title: "Simulations from Time Series Models"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Simulations from Time Series Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
library(simts)
```

In this section, we briefly list, describe, and provide the syntax from the `simts` package that may be useful for creating simulation methods involving time-series models.  

The following list includes some basic models available in the `simts` package.

* White Noise `WN()`
* Quantization Noise `QN()`
* Random Walk `RW()`
* Drift `DR()`
* First-order Autoregressive `AR1()`
* Autoregressive `AR()`
* First-order Moving Average `MA1()`
* Moving Average Process `MA()`
* Gauss-Markov `GM()`
* Autoregressive Moving Average `ARMA()`
* Integrated Autoregressive Moving Average `ARIMA()`
* Seasonal Autoregressive Integrated Moving Average `SARIMA()`
* Seasonal Autoregressive Moving Average `SARMA()`
* Quantization Noise `QN()`

Quantization Noise is a less known process that is used in engineering applications. It can be described in layperson terms as being a good estimator of a rounding error.

## Generate Simulations 

The code below shows how to call the function `gen_gts()`, which allows the user to generate samples from the above model specifications.

```{r}
# Set seed for reproducibility
set.seed(1337)

# Number of observations
n = 1e4

# Generate a White Noise Process
wn = gen_gts(n, WN(sigma2 = 1)) 

# Generate a Quantization Noise
qn = gen_gts(n, QN(q2 = .5)) 

# Generate a Random Walk
rw = gen_gts(n, RW(gamma2 = .75)) 
```

By applying the `plot()` function on the result of a `gen_gts()` simulation, we can observe a visualization of our simulated data.

```{r, fig.align='center', fig.height = 4, fig.width = 7.25, fig.cap = 'Figure 1: Simulated SARIMA(1,0,1)x(2,1,1)[12] process'}
# Generate an SARIMA(1,0,1)x(2,1,1)[12]
sarima = gen_gts(n, SARIMA(ar = 0.3, i = 0, ma = -0.27,
                        sar = c(-0.12, -0.2), si = 1, sma = -0.9,
                        sigma2 = 1.5, s = 12))
# Plot simulation of SARIMA(1,0,1)x(2,1,1)[12]
plot(sarima)
```

The `simts` package therefore allows users to easily simulate from a wide variety of classical time series models, but does not limit itself to these models. Indeed, under some restrictions, these models can be combined in different ways to deliver many state-space (latent) models which can be represented as the sum of basic models.

`simts`'s user friendly interface allows for easy construction of such linear state-space models. In fact, to specify that a certain model is a combination of different models, all that is needed is the “+” symbol between them. 

Suppose that different AR1 processes are present in a state-space model. The syntax to insert “k” of these models into the state-space model is `k*AR1()`. So, for example, the sum of three AR1 models, a random walk and a white noise process can be given by a simple expression: `3*AR1()+RW()+WN()`.

Examples of simulating such models are generated below.

```{r, fig.align='center', fig.height = 4, fig.width = 7.25, fig.cap = 'Figure 2: Simulated ARMA(2,1) + WN() process'}
# Generate a ARMA(2,1) + WN()  
arma_wn_model =
  ARMA(ar = c(0.9, -0.5), ma = 0.3, sigma2 = 1) + 
  WN(sigma = 4)
arma_wn_sim = gen_gts(n = n, model  = arma_wn_model)
# Plot simulation of ARMA(2,1) + WN()
plot(arma_wn_sim)
```

## Latent Time Series

`simts` provides a function specifically designed to generate and represent latent time series models: `gen_lts()`. This provides users the option to visualize a breakdown of the underlying processes by applying the `plot()` function on the result of `gen_lts()`.

```{r, fig.align='center', fig.height = 7, fig.width = 7.25, fig.cap = 'Figure 3: Simulated SARMA(1,0) x (0,1) + WN(2) process with a breakdown of the underlying latent processes'}
# Generate a SARMA() + WN() 
sarma_wn_model = 
  SARMA(ar = 0, ma = 0, sar = 0.9, sma = 0, s = 10, sigma2 = 1) + 
  WN(sigma2 = 2)
sarma_wn_sim = gen_lts(n = n, model = sarma_wn_model)

# Plot simulation of SARMA() + W() 
plot(sarma_wn_sim)
```

