---
author: "Justin Lee"
title: "Simulations from Time Series Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
library(simts)
```

In this section, we briefly list, describe, and provide the syntax from the `simts` package that may be useful for creating simulation methods involving time-series models. 

The following list includes some basic models available in the `simts` package.

* White Noise `WN()`
* Quantization Noise `QN()`
* Random Walk `RW()`
* Drift `DR()`
* First-order Autoregressive `AR1()`
* Create an Autoregressive P `AR()`
* First-order Moving Average `MA1()`
* Moving Average Q Process `MA()`
* Gauss-Markov `GM()`
* Quantisation Noise `QN()`
* Autoregressive Moving Average `ARMA()`
* Integrated Autoregressive Moving Average `ARIMA()`
* Seasonal Autoregressive Integrated Moving Average `SARIMA()`
* Seasonal Autoregressive Moving Average `SARMA()`

Quantization Noise is a lesser known process that is used in engineering applications. It can be described in layperson terms as being a good estimator of a rounding error.

## Simulations

The code below shows how to call the function `gen_gts()`, which allows the user to generate samples from these model specifications.

```{r}
# Set seed for reproducibility
set.seed(1337)

# Number of observations
n = 1e4

# Generate a White Noise Process
wn = gen_gts(n, WN(sigma2 = 1)) 

# Generate a Quantization Noise
qn = gen_gts(n, QN(q2 = .5)) 

# Generate a Random Walk
rw = gen_gts(n, RW(gamma2 = .75)) 

# Generate a Drift
dr = gen_gts(n, DR(omega = 0.10)) 

# Generate an AR(1)
ar1 = gen_gts(n, AR1(phi = .9, sigma2 = .1)) 

# Generate an MA(1)
ma1 = gen_gts(n, MA1(theta = .3, sigma2 = .5)) 

# Generate an ARMA(1,1)
arma11 = gen_gts(n, ARMA11(phi = .9, theta = .2, sigma2 = 1))

# Generate an SARIMA(1,0,1)x(2,1,1)[12] process
sarima = gen_gts(n, SARIMA(ar = 0.3, i = 0, ma = -0.27,
                        sar = c(-0.12, -0.2), si = 1, sma = -0.9,
                        sigma2 = 1.5, s = 12))
```

The simts package therefore allows users to easily simulate from a wide variety of models, but does not limit itself to these basic models. Indeed, under some restrictions, these models can be combined in different ways to deliver many state-space (latent) models which can be represented as the sum of basic models.

`simts`'s user friendly interface allows for easy construction of such linear state-space models. In fact, to specify that a certain model is a combination of different models, all that is needed is the “+” symbol between them. 

Supposing that different AR1 processes are present in the state-space model, the syntax to insert “k” of these models in a state-space model is `k\*AR1()`. So, for example, the sum of three AR1 models, a random walk and a white noise process can be given by a simple expression: `3\*AR1()+RW()+WN()`.

## Visualizations 

By applying the `plot()` function on the result of `gen_gts()`, we can observe a visualization of the simulated data!

```{r, fig.align='center', fig.height = 5, fig.width = 7.25}
# Plot simulation of a SARIMA Model 
plot(sarima)

# Generate a ARMA + WN process 
arma_wn_model =
  ARMA(ar = c(0.9, -0.5), ma = 0.3, sigma2 = 1) + 
  WN(sigma = 4)
arma_wn_sim = gen_gts(n = n, model  = arma_wn_model)
# Plot simulation 
plot(arma_wn_sim)
```

Finally, `simts` provides a function specifically designed to generate and represent latent time series models: `gen_lts()`. This provides users the option to plot a breakdown of the underlying processes by applying the `plot()` function on the result of `gen_lts()`.

```{r, fig.align='center', fig.height = 7, fig.width = 7.25}
sarma_wn_model = 
  SARMA(ar = 0, ma = 0, sar = 0.9, sma = 0, s = 10, sigma2 = 1) + 
  WN(sigma2 = 2)
Xt = gen_lts(n = n, model  = sarma_wn_model)
plot(Xt)

sarma_wn_ma_model = 
  SARMA(ar = 0.25, ma = 0, sar = 0.9, sma = 0, s = 10, sigma2 = 1) + 
  WN(sigma2 = 2) +
  MA(theta = c(0.1,-0.8), sigma2 = 2)
Yt = gen_lts(n = n, model  = sarma_wn_ma_model)
plot(Yt)
```

